Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-03-30 paper/main.tex
fixing up table references

\begin{figure}[hbtp]
\centering
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\hline
\end{tabular}
\end{figure}

\begin{table}[hbtp]
\centering
\caption{Regular Expression Rule Examples}
\label{tab:rules}
\resizebox{\columnwidth}{!}{%

  \begin{tabular}{|l|l|l|}
    \hline
    Universal & C-Like & Language-Specific \\
    \hline
{\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
{\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
                                                                 {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
                 (1==1)} & {\tt any\_of ==> all\_of}\\
\hline

\end{tabular}
}

\begin{table}[hbtp]
\centering
\caption{Parser Parser Combinator Rule Examples}
\label{tab:crules}
\resizebox{\columnwidth}{!}{%

  \begin{tabular}{|l|l|l|}
    \hline
    Universal & C-Like & Language-Specific \\
    \hline
{\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
{\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
                                                                 {\tt
                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
{\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
\hline

\end{tabular}
}
\end{table}

Table~\ref{tab:rules} shows examples of universal, C-like, and
language-specific rules for a few languages, taken directly from our
implementation of regular-expression based operators (see Section \ref{sec:imp}).  In all cases, the form of a rule is: {\tt regexp
  ==> replacement}, using Python regular expression and replacement
syntax.  While these rules are non-trivial to write and read, we show
below that using parser parser combinators often simplifies and
clarifies such rules.

@@

Table~\ref{tab:crules} shows how the rules in Table~\ref{tab:rules}
look, when implemented as Comby patterns.  While some simple rules are
unchanged, others are either more succinct or provide more context as
to applicability (e.g., only less than between valid holes should be
mutated).  This shows that a
further major benefit of using Comby: once a user has learned
the simple concepts above, it is often much easier to read---and
especially \emph{write}---mutation
operator definitions in Comby than using regular expressions.  The
added precision comes not only ``for free'' but in fact yields a
benefit for the ``end-user-programmer'' of mutations.

@@

Our comparison of Python mutation tools showed the largest degree of
similarity between all tools.  Muation scores and equivalence rates
were all, given the large variance expected from previous studies of
Java tools, quite similar.  Again, there is no clear disadvantage to
using our approach, and a large advantage in terms of development and
maintenance ease.  Comby had a somewhat higher equivalence rate than the
other tools (including the Regexp mode of our tool).  We discuss
possible reasons for this below.

>>>>

\begin{figure}[hbtp]
\centering
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\end{figure}

% \begin{table}[hbtp]
% \centering
% \caption{Regular Expression Rule Examples}
% \label{tab:rules}
% \resizebox{\columnwidth}{!}{%

%   \begin{tabular}{|l|l|l|}
%     \hline
%     Universal & C-Like & Language-Specific \\
%     \hline
% {\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
% {\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
%                                                                  {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
% {\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
%                  (1==1)} & {\tt any\_of ==> all\_of}\\
% \hline

% \end{tabular}
% }

% \end{table}

% \begin{table}[hbtp]
% \centering
% \caption{Parser Parser Combinator Rule Examples}
% \label{tab:crules}
% \resizebox{\columnwidth}{!}{%

%   \begin{tabular}{|l|l|l|}
%     \hline
%     Universal & C-Like & Language-Specific \\
%     \hline
% {\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
% {\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
%                                                                  {\tt
%                                                                           :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
% {\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
% \hline

% \end{tabular}
% }
% \end{table}

The left-hand-side Table~\ref{tab:rules} shows examples of universal, C-like, and
language-specific rules for a few languages, taken directly from our
implementation of regular-expression based operators (see Section \ref{sec:imp}).  In all cases, the form of a rule is: {\tt regexp
  ==> replacement}, using Python regular expression and replacement
syntax.  While these rules are non-trivial to write and read, we show
below that using parser parser combinators often simplifies and
clarifies such rules.

@@

The right-hand-side of Table~\ref{tab:rules} shows the same regular-expression rules 
implemented as Comby patterns.  While some simple rules are
unchanged, others are either more succinct or provide more context as
to applicability (e.g., only less than between valid holes should be
mutated).  This shows that a
further major benefit of using Comby: once a user has learned
the simple concepts above, it is often much easier to read---and
especially \emph{write}---mutation
operator definitions in Comby than using regular expressions.  The
added precision comes not only ``for free'' but in fact yields a
benefit for the ``end-user-programmer'' of mutations.

@@

Our comparison of Python mutation tools showed the largest degree of
similarity between all tools.  Mutation scores and equivalence rates
were all, given the large variance expected from previous studies of
Java tools, quite similar.  Again, there is no clear disadvantage to
using our approach, and a large advantage in terms of development and
maintenance ease.  Comby had a somewhat higher equivalence rate than the
other tools (including the Regexp mode of our tool).  We discuss
possible reasons for this below.
